---
title: 'The Study of ''Other'': How Athletes and Students of Different Ethnicities
  Perform Academically at Williams'
author: "Haoyu Sheng"
date: "December 7, 2017"
output:
  slidy_presentation: 
    incremental: true
---
  
```{r setup, echo=FALSE, include=FALSE}
require(tidyverse) #you may need to run install.packages('tidyverse') first, only once on a machine
require(dplyr)
require(lubridate) #same goes for any new package
require(rvest)
require(stringr)
require(knitr)
require(ggplot2)
load("../data/honors.RData")
load(file = "../data/athletic_summary.RData")
require(gridExtra)
load("../data/students.RData")
load("../data/roster.RData")
load("../data/classified.RData")
load(file = "../data/ethnic_summary.RData")

knitr::opts_chunk$set(tidy=TRUE, message=FALSE)
```


#Motivation and Approaches

- **Motivation**

    - Informal Response to the article : *What Does it Mean to be the Best?: An Alum Considers the Relative Importance of Admission Criteria*

    - Personal Interest in seeing how students of Asian origins are performing academically

- **Approach**

    - Attain information from graduation catalog and team rosters

    - Use the `wru` package to identify someone's ethnicity through surname analysis

    - Compare the percentage of students attaining `cum laude` among different demographics. 

----------------------

#Disclaimers and Limitations

- Limitations of the Bayesian surname analysis. Inability to account for names such as "Lee"
- Loss of information during transition. For example: a student named Andrew might be recorded as Andy on the team roster
- Inability to account for walk-ons.
- Graduation honors are not the only metrics for a student's importance and contribution to the college!

-----------------------

#Data 

- Utilizes two datasets: **student** and **roster**
- The `students` dataset comes from the Williams College catalog archive, containing graduation data from the class of 2003 to the class of 2015
- Has `r nrow(students)` entries

    ```{r, echo = FALSE}
    head(students, 3) %>%
      mutate(name = c('*Amy Eph, with highest honor in Economics', '*Chris Somebody', '*Lauren Nobody')) %>%
      kable()
    ```

- The `roster` dataset comes forom Williams College athletic archives, including rosters from the class of 2002 to the class of 2015 and teams such as soccer, football, and cross country
- Has `r nrow(roster)` entries
    ```{r, echo = FALSE}
    head(roster, 3) %>%
      mutate(name = c('John Lax', 'Ryan Tennis', 'Will Football')) %>%
      kable()
    ```


----------

#Data Tidying and Manuvering

- Join the two initial dataframes
- Read honors based on the names
- Analyze ethnicities based on surnames
- summarize ratio for each subgroup for each year
- 
    ```{r, echo = FALSE}
    kable(head(ethnic_summary, 3))
    ```
    
- A graphical summary of the data
    
    ```{r, echo=FALSE, message = FALSE, warning = FALSE}
    p1 <- ethnic_summary %>%
      ggplot(aes(x = year, y = ratio, color = race)) + geom_point() + theme_bw() 
    p2 <- honors %>%
      mutate(latin = ifelse(summa == 1 | magna == 1 | cum == 1, 1, 0)) %>% 
      group_by(year, team) %>%
      summarize(ratio = sum(latin) / length(name)) %>%
      filter(team == "Women's Cross Country" |  team == "Men's Soccer" | team == "football" | is.na(team)) %>%
      ggplot(aes(x = year, y = ratio, color = team)) + geom_point() + theme_bw() 
    grid.arrange(p1, p2)
    ```

----------------------

#Two proportion z-test

-
```{r}
hispanic <- classified %>%
  filter(hispanic == 1) %>%
  group_by(latin) %>%
  summarize(value = length(name))
non_hispanic = classified %>%
  filter(hispanic == 0) %>%
  group_by(latin) %>%
  summarize(value = length(name))
prop.test(c(filter(hispanic, latin == 1)$value, 
            filter(non_hispanic, latin == 1)$value), 
          c(sum(hispanic$value), sum(non_hispanic$value)))
```
- Independence
- Success/Failure Condition
- 10% Condition


---------------------------

#T tests

```{r}
athlete <- athletic_summary %>%
  filter(athlete == 1) 
non_athlete <- athletic_summary %>%
  filter(athlete == 0) 
t.test(athlete$ratio, non_athlete$ratio)
```

- Conditions
- Nearly Normal: 
    ```{r, echo = FALSE}
    p1 <- ggplot(athlete, aes(x = ratio, fill = "athlete")) + 
      geom_histogram(bins = 6, aes(y = ..density..)) + theme_bw() + 
      stat_function(fun = dnorm, 
                    args = c(mean = mean(athlete$ratio), 
                             sd = sd(athlete$ratio)))
    p2 <- ggplot(non_athlete, aes(x = ratio, fill = "non-athlete")) + 
      geom_histogram(bins = 6, aes(y = ..density..)) + theme_bw() +
      stat_function(fun = dnorm, 
                    args = c(mean = mean(non_athlete$ratio), 
                             sd = sd(non_athlete$ratio)))
    grid.arrange(p1, p2, nrow = 1)
    ```
- Independence


-------------------------

#Thoughts and Potential Future Work

- Different subgroups of students might have different focuses
- Interesting data for the Davis center
- A foundation for a study on the relationship between socioeconomic status and thriving academically at college
- Compare results across colleges to examine if our institution is doing its best to support students of various demographics



